\chapter{Summary}
Document presents comparison between different C++ Finite State Machine frameworks.
Tested frameworks vary in their terms of functionality, approach and concept.
Frameworks were compared due to compilation speed, execution time, debug size, strip size, memory
consumption, functionality and UML compatibility. Also different compilers and compilation flags were taken into account.
Results shows that declarative concept might be as efficient as imperative concept
while preserving at the same transparency of the state machine, as well as easy to test
code. Tests also indicate that there is no one way to implement FSM framework which meet all the
given criteria. Therefore frameworks like msm, QFsm have the advantage that the policy may be
adjusted to better meet some of the requirements.
In most cases compilation time seems to be longer in proportion to execution time.
This does not apply to QFsm.FavorCompilationTime for which execution speed is kept in most tests as well
as compilation time. QFsm.FavorExecutionSpeed and msm in case of compilation time seems to be far away after opponents.
Execution speed shows that msm, QFsm.FavorExecutionSpeed, QFsm.FavorCompilationTime are the fastest.
StateChart, QFsm.FavorDebugSize are about 10 times slower.
Debug size is huge in msm and QFsm.FavorExecutionSpeed and the remaining frameworks keep the debug size
within reasonable limits. Release (strip) size is the smallest in QFsm.FavorCompilationTime.
About 2 times bigger in QFsm.FavorExecutionSpeed, 3 times bigger in msm, statechart.
Memory consumption meets the requirements in all frameworks besides the statechart
framework.
